---
title: "501 Final Analysis"
output: html_document
---

```{r}
library(tidyverse)
library(infer)
library(readr)
library(tidyr)
library(lme4)
library(readxl)

library(BayesFactor)
library(dplyr)
library(ggplot2)
library(cowplot)
library(grid)
library(gridExtra)
library(plotrix)
library(ggridges)
library(ggpubr)
library(stargazer)
```
```{r}
#Load data
data <- read_excel("Group8_Data_NumericValues_clean.xls", col_names = FALSE, skip = 3)
varnames <- read_excel("Group8_Data_NumericValues_clean.xls", n_max = 1)
colnames(data) <- colnames(varnames)
```
#### Importing and cleaning dataset

```{r}
#"Group8_Data_NumericValues_clean.xls" contains only responses which have passed the qualitative exclusion criteria for T1 and T2

#Identify columns of interest 
clean <- data %>% select(6, 24:27, 32, 37, 43:52, 58:67,72:75, 91, 92, 96) %>% rename(duration = `Duration (in seconds)`)
```
```{r}
#Primary and secondary outcome variable are calculated here
#Time based exclusion criteria are filtered here. Speeders are removed
clean1 <- clean %>% filter(duration >= 129) %>%  mutate(conditionLabel= case_when(
                                     Condition == 1 ~ "Past- No Control",
                                     Condition == 2 ~ "Past- In Control",
                                     Condition == 3 ~ "Present- No Control (Ultimatum Game)",
                                     Condition == 4 ~ "Present- In Control (Dictator Game)")) %>% 
                                       rowwise() %>% 
                                                mutate( meanPreDV =  mean(preDV1, preDV2),
                                                      meanDV = mean(DV1, DV2),
                                                      meanPreC = mean(preC1, preC2),
                                                      meanC = mean(C1, C2),
                                                      deltaC = meanPreC-meanC
                                                      ) 
```

```{r}
# Function for Bayes factor
BFplus0 <- function(BFmodel,BFeffect, test.rule){
  BFmodel      <- BFmodel
  BFeffect     <- BFeffect 
  consistent   <- test.rule
  N_consistent <- sum(consistent)
  bf_re        <- (N_consistent / 10000) / (1 / 2)
  bf_e0        <- as.vector(BFeffect)
  ## Use transitivity to compute desired Bayes factor
  bf_r0        <- bf_re * bf_e0
  bf_plus_0    <- bf_r0
  return(bf_plus_0)} 

```

#### Distribution of OV Scores
```{r}
#Plotting the distribution of Belief in God Scores
plot_labels <- c("1" = "T1", "2" = "T2", "3" = "T3", "4" = "T4")
ggplot(clean1, aes(x = meanDV)) + facet_wrap(Condition ~ ., labeller = labeller(Condition = plot_labels)) + geom_histogram()  + labs(x= "Belief in God Score", y= "Number of Respondents") + scale_x_continuous(breaks = seq(1, 7, 1), lim = c(1, 7)) +theme_bw() 

```

#### Hypothesis 1 Analysis
H1: Perceptions about lack of control in the present will lead to an increased belief in the existence of a controlling God
```{r}
# Recoding conditions for hypothesis testing T3 vs Others
H1recode            <- clean1 %>% mutate(h1comp = case_when(
                                     Condition == 1~"null" ,
                                     Condition == 2~"null" ,
                                     Condition == 3~"alt" ,
                                     Condition == 4~"null" ))
H1null              <- lmBF(meanDV ~ Reli, data = H1recode)
H1alt               <- lmBF(meanDV ~ h1comp + Reli, data = H1recode)
BFH1                <- H1alt/H1null
samples             <- posterior(H1alt, iterations = 10000)

mean.dif.condition  <- samples[,"h1comp-alt"] - samples[,"h1comp-null"]
med.mean.dif.con    <- median(mean.dif.condition)
grant.slope.rel     <- samples[,"Reli-Reli"]
med.grant.slope.rel <- median(grant.slope.rel)
test.rule           <- (samples[,"h1comp-alt"] < samples[,"h1comp-null"])
BF_rest_exp         <- BFplus0(H1alt, BFH1, test.rule)

#Interaction effect model exploration
Mfull               <- lmBF(meanDV ~ Reli + h1comp + Reli:h1comp, data = H1recode)
Mint                <- Mfull/H1alt

# Calculate prior and posterior model probabilities 
BFs            <- c(H1null, H1alt, Mfull)           #get relevant models to compare 
prior.odds     <- newPriorOdds(BFs, type = "equal") #set all prior odds to 1
post.odds      <- prior.odds * BFs                  #calculate posterior odds based on Bayes' rule
post.prob      <- as.BFprobability(post.odds)       #Normalize to add up to 1 (probabilities)
posterior.prob <- post.prob
posterior.prob                                      # analyzed output
```
```{r}
#Calculate Wilcoxon Rank Sum median differences.
H1test <- compare_means(meanDV ~ h1comp, data= H1recode, method = "wilcox.test", adjustment = "holmes")
H1test
```
```{r}
# Create boxplot visualization with Wilcoxon p-value
comparisons <- list( c("null", "alt") )
H1recode %>%      ggplot(aes(x= h1comp, y= meanDV)) + 
                  geom_boxplot() + 
                  stat_compare_means(comparisons= comparisons) + 
                  labs(x= "Condition", y= "Belief in God") + 
                  geom_jitter(width = 0.2, height = 0.2) + 
                  scale_y_continuous(breaks = seq(0, 8, 1), lim = c(0, 8)) + 
                  theme_classic() + 
                  scale_x_discrete(labels= c( "Present- No Control", "Null")) 
```
#### Hypothesis 2 Analysis
H2: Situations in which one had no personal control will result in a more fervent belief in the existence of a controlling God, compared to situations in which one does have personal control.
```{r}
# Recoding conditions for hypothesis testing no-control(T1 and T3) vs in-control (T2 and T4)
H2recode            <- clean1 %>% mutate(h2var = case_when(
                                     Condition == 1~"no" ,
                                     Condition == 2~"in" ,
                                     Condition == 3~"no" ,
                                     Condition == 4~"in" ))
H2null              <- lmBF(meanDV ~ Reli, data = H2recode)
H2alt               <- lmBF(meanDV ~ h2var + Reli, data = H2recode)
BFH2                <- H2alt/H2null
samples2            <- posterior(H2alt, iterations = 10000)

mean.dif.condition2  <- samples2[,"h2var-no"] - samples2[,"h2var-in"]
med.mean.dif.con2    <- median(mean.dif.condition2)
grant.slope.rel2     <- samples2[,"Reli-Reli"]
med.grant.slope.rel2 <- median(grant.slope.rel2)
test.rule2           <- (samples2[,"h2var-no"] < samples2[,"h2var-in"])
BF_rest_exp2         <- BFplus0(H2alt, BFH2, test.rule2)

#Interaction effect model exploration
M2full              <- lmBF(meanDV ~ Reli + h2var + Reli:h2var, data = H2recode)
M2int               <- M2full/H2alt

# Calculate prior and posterior model probabilities
BFs2            <- c(H2null, H2alt, M2full)           #get relevant models to compare 
prior.odds2     <- newPriorOdds(BFs2, type = "equal") #set all prior odds to 1
post.odds2      <- prior.odds2 * BFs2                 #calculate posterior odds based on Bayes' rule
post.prob2      <- as.BFprobability(post.odds2)       #Normalize to add up to 1 (probabilities)
posterior.prob2 <- post.prob2
posterior.prob2                                       # analyzed output

```
```{r}
#Calculate Wilcoxon Rank Sum median differences.
H2test <- compare_means(meanDV ~ h2var, data= H2recode, method = "wilcox.test", adjustment = "holmes")
H2test
```
```{r}
# Create boxplot visualization with Wilcoxon p-value
H2recode %>%      ggplot(aes(x= h2var, y= meanDV)) + 
                  geom_boxplot() + 
                  stat_compare_means(comparisons = list( c("in", "no") )) + 
                  labs(x= "Condition", y= "Belief in God") + 
                  geom_jitter(width = 0.2, height = 0.2) + 
                  scale_y_continuous(breaks = seq(0, 8, 1), lim = c(0, 8))  + 
                  theme_classic() + 
                  scale_x_discrete(labels= c( "In-Control", "No-Control"))
```
#### Hypothesis 3
H3:  Engaging in the Ultimatum Game will be more effective in inducing a lowered sense of control as compared to autobiographical recall. 
```{r}
# Recoding conditions for hypothesis testing T1 vs T3 no-control conditions
H3recode              <- clean1 %>% mutate(h3var = case_when(
                                     Condition == 1~"no" ,
                                     Condition == 2~"in" ,
                                     Condition == 3~"no" ,
                                     Condition == 4~"in" )) %>% filter(h3var== "no") %>% 
                       mutate(temporality =  case_when(
                                       Condition ==1 ~ "past",
                                       Condition == 3 ~ "present"))
H3null                <- lmBF(deltaC ~ Reli, data = H3recode)
H3alt                 <- lmBF(deltaC ~ temporality + Reli, data = H3recode)
BFH3                  <- H3alt/H3null
samples3              <- posterior(H3alt, iterations = 10000)

mean.dif.condition3  <- samples3[,"temporality-past"] - samples3[,"temporality-present"]
med.mean.dif.con3    <- median(mean.dif.condition3)
grant.slope.rel3     <- samples3[,"Reli-Reli"]
med.grant.slope.rel3 <- median(grant.slope.rel3)
test.rule3           <- (samples3[,"temporality-past"] < samples3[,"temporality-present"])
BF_rest_exp3         <- BFplus0(H3alt, BFH3, test.rule3)

#Interaction effect model exploration
M3full                <- lmBF(deltaC ~ Reli + temporality + Reli:temporality, data = H3recode)
M3int                 <- M3full/H3alt

# Calculate prior and posterior model probabilities 
BFs3                  <- c(H3null, H3alt, M3full)           #get relevant models to compare 
prior.odds3           <- newPriorOdds(BFs3, type = "equal") #set all prior odds to 1
post.odds3            <- prior.odds3 * BFs3                 #calculate posterior odds based on Bayes' rule
post.prob3            <- as.BFprobability(post.odds3)       #Normalize to add up to 1 (probabilities)
posterior.prob3       <- post.prob3
posterior.prob3

```
```{r}
#Calculate Wilcoxon Rank Sum median differences.
compare_means(deltaC ~ temporality, data= H3recode, method = "wilcox.test", adjustment = "holmes")
```
```{r}
# Create boxplot visualization with Wilcoxon p-value
H3recode %>%          ggplot(aes(x= temporality, y= deltaC)) + 
                      geom_boxplot() + 
                      stat_compare_means(comparisons= list( c("present", "past") )) + 
                      labs(x= "Condition", y= "Change in Sense of Control") + 
                      geom_jitter(width = 0.2, height = 0.2) + 
                      scale_y_continuous(breaks = seq(-6, 6, 1), lim = c(-6, 6))  + 
                      theme_classic() + 
                      scale_x_discrete(labels= c( "Past", "Present"))
```
#### Hypothesis 4
H4: Engaging in the Dictator Game will be more effective in inducing an increased sense of control as compared to autobiographical recall.
```{r}
# Recoding conditions for hypothesis testing T2 vs T4 in-control conditions
H4recode      <- clean1 %>% mutate(h4var = case_when(
                                     Condition == 1~"no" ,
                                     Condition == 2~"in" ,
                                     Condition == 3~"no" ,
                                     Condition == 4~"in" )) %>% filter(h4var== "in") %>% 
                      mutate(temporality =  case_when(
                                       Condition ==2 ~ "past",
                                       Condition == 4 ~ "present"))

H4null                <- lmBF(deltaC ~ Reli, data = H4recode)
H4alt                 <- lmBF(deltaC ~ temporality + Reli, data = H4recode)
BFH4                  <- H4alt/H4null
samples4              <- posterior(H4alt, iterations = 10000)

mean.dif.condition4  <- samples4[,"temporality-past"] - samples4[,"temporality-present"]
med.mean.dif.con4    <- median(mean.dif.condition4)
grant.slope.rel4     <- samples4[,"Reli-Reli"]
med.grant.slope.rel4 <- median(grant.slope.rel4)
test.rule4           <- (samples4[,"temporality-past"] < samples4[,"temporality-present"])
BF_rest_exp4         <- BFplus0(H4alt, BFH4, test.rule4)

#Interaction effect model exploration
M4full <- lmBF(deltaC ~ Reli + temporality + Reli:temporality, data = H4recode)
M4int  <- M4full/H4alt

# Calculate prior and posterior model probabilities 
BFs4            <- c(H4null, H4alt, M4full)           #get relevant models to compare 
prior.odds4     <- newPriorOdds(BFs4, type = "equal") #set all prior odds to 1
post.odds4      <- prior.odds4 * BFs4                 #calculate posterior odds based on Bayes' rule
post.prob4     <- as.BFprobability(post.odds4)        #Normalize to add up to 1 (probabilities)
posterior.prob4 <- post.prob4
posterior.prob4

```
```{r}
#Calculate Wilcoxon Rank Sum median differences.
compare_means(deltaC ~ temporality, data= H4recode, method = "wilcox.test", adjustment = "holmes")
```
```{r}
# Create boxplot visualization with Wilcoxon p-value
H4recode %>%              ggplot(aes(x= temporality, y= deltaC), color= (temporality)) + 
                          geom_boxplot() + 
                          stat_compare_means(comparisons= list( c("present", "past"))) + 
                          labs(x= "Condition", y= "Change in Sense of Control") + 
                          geom_jitter(width = 0.2, height = 0.2) + 
                          scale_y_continuous(breaks = seq(-4, 4, 1), lim = c(-4, 4)) + 
                          theme_classic() + 
                          scale_x_discrete(labels= c( "Past", "Present"))
```

#### Between Group Analysis
```{r}
# Computing Wilcoxon Rank Sum tests across all condition pairs
compare_means(meanDV ~ (Condition), data= clean1, method = "wilcox.test", adjustment = "holmes")
```

```{r}
#Plotting Belief in God across all conditions with pairwise Wilcoxon p-values for comparable treatments
comparisons <- list( c("4", "3"), c("4", "2"), c("3", "1"), c("2", "1") )
clean1 %>%            ggplot(aes(x= factor(Condition), y= meanDV, color= factor(Condition))) + 
                      geom_boxplot() + stat_compare_means(comparisons= comparisons) + 
                      labs(x= "Condition", y= "Belief in God") + 
                      scale_x_discrete(labels= c("Past- No Control", "Past- In Control", "Present- No Control", "Present- In Control")) + 
                      geom_jitter(width = 0.3, height = 0.3) + 
                      scale_y_continuous(breaks = seq(0, 8, 1), lim = c(0, 10)) + 
                      theme_classic()
```

```{r}
#Compute Kruskal Wallis median comparison test 
compare_means(meanDV ~ Condition, data= clean1, method= "kruskal.test", adjustment = "holmes")
```
#### OLS Robustness Check
```{r}
# Model to check for significance across conditions and covariates 
olscheck <- (lm(meanDV ~  factor(Condition) + Reli + meanC, data = clean1))
summary(olscheck)
```
```{r mylatextable, results = "asis", echo=FALSE}
#Formatting output table for the OLS check
stargazer(olscheck, type = 'latex', 
          title = 'Regression table of OLS regression', 
          single.row=TRUE)
```

#### Appendix L calculations

```{r}
# Means and Standard Deviations by condition
meansAndSDs <- clean1 %>% select(Condition, meanDV, meanC) %>% group_by(Condition) %>% summarize( meanGod = mean(meanDV), sdGod = sd(meanDV), meanCon = mean(meanC), sdCOn = sd(meanC))
meansAndSDs
```






